---
title: "PMLProject"
author: "Sammy kariuki"
date: "October 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Excutive Summary   

Based on a dataset provided by HAR [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har) we will train various predictive models and select the best and most accurate model to predict what exercise was performed using the provided pml-training dataset with 159 features.    

We will take the following steps:

- Load and process the raw data    
- 


#### Load libraries   
    
This section we load the R package that we shall use in this Project.    

```{r, warning=FALSE, comment= FALSE, message=FALSE}
library(readr)

library(caret)

library(rattle)

library(randomForest)
```

#### Load Data   
   
```{r, warning=FALSE, comment= FALSE, message=FALSE}
pml_training <- read_csv("./pml-training.csv")

pml_testing<- read_csv("./pml-testing.csv")
```
       
## Preprocessing
       
#### Select on the relevant variable and create training and validation dataset   
    
```{r, warning=FALSE, comment= FALSE, message=FALSE}
# relevant variable by the proportion missing data

# 1. Training

relevant_var <- names(pml_training)[(colSums(is.na(pml_training))/dim(pml_training)[1] < 0.50)]

# 2. Testing

relevant_vartest <- names(pml_testing)[(colSums(is.na(pml_testing))/dim(pml_testing)[1] < 0.50)]

# Check if there is any variable relevant in training and not in testing data set

"%ni%" <- Negate("%in%")
relevant_var[relevant_vartest  %ni% relevant_var]
```
      
    
    
```{r, warning=FALSE, comment= FALSE, message=FALSE}
# select relevant variable

pml_test <- pml_testing[, relevant_vartest]

pml_test <- pml_test[, -c(1:7)]

pml_train <- pml_training[, relevant_var]

pml_train <- pml_train[, -c(1:7)]

row_has_NA <- apply(pml_train, 1, function(x){any(is.na(x))})

pml_train <- pml_train[!row_has_NA, ]

pml_train$classe <- as.factor(pml_train$classe)


# partition train and validataion set

set.seed(5000)

InTrain <- createDataPartition(pml_train$classe, p = 0.7, list = FALSE)

training <- pml_train[InTrain, ]

testing <- pml_train[-InTrain, ]
```
     
#### Exploratory Analysis

```{r, warning=FALSE, comment= FALSE, message=FALSE}
# Compare dimension of data set before and after eliminating less relevant variable

dim(pml_training) # original available dataset

dim(training) # used training model

dim(testing) # used for varidation of model

dim(pml_testing) # unseen data to predict

# check if the data is biased

table(training$classe)

table(testing$classe)
```
     
#### Fit a model rpart in caret package   
     
```{r, warning=FALSE, comment= FALSE, message=FALSE}
RpartModel_fit <- train(classe ~., method = "rpart", data = training)

Validation_pred <- predict(RpartModel_fit, testing)

confusionMatrix(Validation_pred, testing$classe)$overall[1]

fancyRpartPlot(RpartModel_fit$finalModel)
```
      
#### fit random forest model    
     
```{r, warning=FALSE, comment= FALSE, message=FALSE}
RFmodel_Fit <- randomForest(classe ~ ., data = training)

pred_valid <- predict(RFmodel_Fit, testing)

confusionMatrix(pred_valid, testing$classe)$overall[1]

RFmodel_Fit$confusion

pred_classe <- predict(RFmodel_Fit, pml_test)

pred_classe
```

